---
title: "\\vspace{-3.0cm} Regression models course project"
author: "Georgy Makarov"
date: "April 29, 2020"
geometry: "left=2cm, right=2cm, top=2cm, bottom=2cm"
output: pdf_document
---

\vspace{-1.5cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Executive summary

This report explores *Motor Trend* magazine's data set of a collection of cars 
in order to find out, which is better for fuel consumption - manual transmission 
or automatic transmission. Cars with manual transmission on average have 7.2 
higher *MPG* than cars with automatic. Transmission type explains 36% of 
variance in *MPG*. Other variables like number of cylinders, power and weight 
are more important to *MPG* than the transmission. The real difference in 
transmission types is only 1.8 MPG.

## Exploratory data analysis

The data set contains 32 observations of 11 variables. All variables are 
numeric. There are five variables, which are better to consider as factor 
variables: *cyl, vs, am, gear, carb*. The distribution of cars by *MPG* is right 
skewed. The majority of cars has fuel consumption between 15 and 20 MPG. The 
cars are mostly represented by 8-cylinder cars with automatic transmission. A car 
transmission usually has 3 gears. Supporting information is in the appendix in 
the *Figure 1*.

```{r data summary}
data(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
mtcars$gear <- as.factor(mtcars$gear)
mtcars$carb <- as.factor(mtcars$carb)
summary(mtcars$mpg)
```

Cars with manual transmission consume less fuel. The median of *MPG* for manual
transmission is around 23, while for automatic transmission it is around 17. The
reference is in the *Figure 2*.

## Regression models

*H~0~*: automatic and manual transimissions *MPG* are the same. We test the *H~0~*
with a t-test. *P-value* of the test is 0.001, confidence interval does not 
contain zero - there is enough evidence to reject *H~0~*.

```{r ttest}
at <- mtcars[mtcars$am == 0,]
mt <- mtcars[mtcars$am == 1,]
t.test(at$mpg, mt$mpg)$p.value
t.test(at$mpg, mt$mpg)$conf
```

Linear regression model with *am* as a predictor quantifies the difference 
between transmissions. *P-values* are significant - average *MPG* for automatic
transmission is 17.1, for manual - 7.2 higher than the automatic. The model
explains 36% of the variance in *MPG* as the *R^2^* is 0.36. We need more 
variables to explain the variance.

```{r regr model}
fit1 <- lm(mpg ~ am, data = mtcars)
summary(fit1)$coeff
summary(fit1)$r.squared
```

One way to choose top predictors is using the *step* function. Top predictors 
are *cyl, hp, wt, am*.

```{r choose predictos}
fit_full <- lm(mpg ~ ., data = mtcars)
step_fit <- step(fit_full, trace = 0)
summary(step_fit)$coeff
```

*P-values* for **multivariate regression** are statistically significant. The
*R^2^* for this model is 0.866. This model shows that manual transmission improves
fuel consumption by 1.81 MPG, while increase in number of cylinders from 4 to 6
causes 3.0 MPG loss. Engines with 8 cylinders eat another 2.2 MPG. Every 
additional *hp* results in decreasing the MPG by 0.03. Every 1000 lbs of weight 
decrease the MPG by 2.5 US gallons.

```{r multivariate regr}
fit4 <- lm(mpg ~ am + cyl + hp + wt, data = mtcars)
summary(fit4)$r.squared
```

The model with four predictors fits better than the model with transmission 
type only. The *anova* test shows the significance of *p-value*.

```{r choose model}
anova(fit1, fit4)
```

## Residuals and diagnostics

Residuals check and model diagnostics are in the *Figure 3*. *Residual vs Fitted* 
plot shows that the residuals are independent. *Normal Q-Q* plot shows normal
distribution of residuals. Residuals on *Scale-Location* plot are randomly 
distributed. There are no outliers as the dots on *Residuals vs Leverage* plot
are within [-0.5;+0.5] interval.

\pagebreak

## Appendix

```{r, echo=FALSE, fig.align="center", fig.width=6, fig.height=6, fig.cap="Distribution"}
par(mfrow = c(2, 2))
hist(mtcars$mpg, main = "Histogram of mpg", col = "skyblue", xlab = "MPG")
barplot(table(mtcars$cyl), main = "Histogram of number of cylinders",
        xlab = "Number of cylinders", ylab = "Frequency", col = "skyblue")
barplot(table(mtcars$am), main = "Histogram of transmission type", 
        col = "skyblue", xlab = "Transmission type", ylab = "Frequency")
barplot(table(mtcars$gear), main = "Histogram of gears", 
        col = "skyblue", xlab = "Number of gears", ylab = "Frequency")
```

```{r, echo=FALSE, fig.align="center", fig.width=6, fig.height=6, fig.cap="Transmission type"}
par(mfrow = c(1, 1))
boxplot(mpg ~ am, data = mtcars, xlab = "Transmission type", col = "skyblue")
```

```{r, echo=FALSE, , fig.align="center", fig.width=6, fig.height=6, fig.cap="Residuals"}
par(mfrow = c(2, 2))
plot(fit4)
```







